### 2020-02-05 paper.tex
Update acknowledgments

\documentclass[fleqn,10pt,twocolumn]{wlscirep}

% Packages
\usepackage{pifont}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{float}
% \usepackage[multiple]{footmisc}
\usepackage{booktabs}
\usepackage{csquotes}
\usepackage{moreverb} % for verbatim ouput (for wordcount)
\usepackage{listings, textcomp}
\lstset{language=Python}

\lstset{ %
  basicstyle=\ttfamily\footnotesize,  % size of fonts used for the code
  xleftmargin=\parindent,
  breaklines=true,   % automatic line breaking only at whitespace
  captionpos=b,   % sets the caption-position to bottom
  commentstyle=\color{gray},  % comment style
  keywordstyle=\color{blue},  % keyword style
  stringstyle=\color{red},  % string literal style
  upquote=true  %straight single quotes (requires textcomp)
}

\usepackage{caption}
\captionsetup{%
   figurename=Fig.,
   tablename=Tab.
}
\usepackage{subcaption}

% New Commands
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\fixme}[1]{\textcolor{red}{{#1}}}
\newcommand{\inlinecite}[1]{\footnotesize\citen{#1}}
\newcommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage[switch]{lineno}
\linenumbers

\title{NumPy---Array Programming in Python}

\input{authors}

\keywords{Scientific computing, Python, Mathematics, Array Programming}

\begin{abstract}
\input{summary} 
\end{abstract}

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

% Count of words
\verbatiminput{wordcount.tex}

\input{body}

\bibliography{references}

\newpage

\input{methods}

\section*{Acknowledgments}

Jim Hugunin wrote Numeric in 1995, while a graduate student at MIT.
Hugunin based his package on previous work by Jim Fulton, then working at the
US Geological Survey, with input from many others.
After he graduated, Paul Dubois at the Lawrence Livermore National Laboratory
became the maintainer.
Many people contributed to the project including Travis Oliphant, David Ascher,
Tim Peters, and Konrad Hinsen.

Around 1998, the Space Telescope Science Institute (STScI) software group began
to use Python heavily, and, around 2000, wrote a reimplementation of
Numeric, called Numarray.
%https://lancesimms.com/programs/Python/pyraf/Docs/greenfield.pdf
Perry Greenfield created the original framework.
Many people contributed to it including Jay Todd Miller, JT Hsu, Richard
L White, Jochen Krupper, and Phil Hodge.

Eric Jones co-founded the SciPy community, gave early feedback on array
implementations, and provided funding and travel support to several community
members.
Numerous people contributed to the creation and growth of the larger SciPy
ecosystem, which gives NumPy much of its value.  Others have injected new
energy and ideas by creating experimental array packages.

Finally, we would like to thank the many members of the community who have provided
feedback, submitted bug reports, made improvements to the documentation,
code, or website, promoted its use in their scientific fields, and built
the vast ecosystem of tools and libraries around NumPy.
Ross Barnowski and Michael Eickenberg suggested text and provided
helpful feedback on the manuscript.

S.J.v.d.W and K.J.M. were funded in part by the Gordon and Betty Moore
Foundation through Grant GBMF3834 and by the Alfred P. Sloan Foundation through
Grant 2013-10-27 to the University of California, Berkeley.
S.J.v.d.W, S.B., M.P., and W.W. were funded in part by the Gordon
and Betty Moore Foundation through Grant GBMF5447 and by the Alfred
P. Sloan Foundation through Grant G-2017-9960 to the University of
California, Berkeley.

\section*{Author Contributions Statement}

K.J.M. and S.J.v.d.W composed the manuscript with input from others.
S.B., R.G., K.S., W.W., M.B., and T.J.R. contributed text.
All authors have contributed significant code, documentation, and/or expertise
to the NumPy project.
All authors reviewed the manuscript.

\section*{Competing Interests}

The authors declare no competing interests.

\end{document}