### 2020-05-25 supplementary.tex

\documentclass[fleqn,10pt,twocolumn]{wlscirep}

% Packages
\usepackage{listings, textcomp}
\lstset{language=Python}

\lstset{ %
  basicstyle=\ttfamily\footnotesize,  % size of fonts used for the code
  xleftmargin=\parindent,
  breaklines=true,   % automatic line breaking only at whitespace
  captionpos=b,   % sets the caption-position to bottom
  commentstyle=\color{gray},  % comment style
  keywordstyle=\color{blue},  % keyword style
  stringstyle=\color{red},  % string literal style
  upquote=true  %straight single quotes (requires textcomp)
}

\usepackage{caption}
\captionsetup{%
   figurename=Fig.,
   tablename=Tab.
}

\usepackage[switch]{lineno}
\linenumbers

\title{Supplementary Methods}

\input{authors}

\begin{abstract}
\input{summary}
\end{abstract}

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\input{methods}

\bibliography{references}

\end{document}