### 2020-08-12 authors.tex
Update pv affil

\author[1]{Charles R. Harris}
\author[2,3,4,*]{K. Jarrod Millman}
\author[5,2,4,*]{St\'efan J. van der Walt}
\author[6,*]{Ralf Gommers}
\author[7,8]{Pauli Virtanen}
\author[9]{David Cournapeau}
\author[10]{Eric Wieser}
\author[11]{Julian Taylor}
\author[4]{Sebastian Berg}
\author[12]{Nathaniel J. Smith}
\author[13]{Robert Kern}
\author[4]{Matti Picus}
\author[14]{Stephan Hoyer}
\author[15]{Marten H. van Kerkwijk}
\author[2,16]{Matthew Brett}
\author[17]{Allan Haldane}
\author[18]{Jaime Fern\'andez del R\'io}
\author[19,20]{Mark Wiebe}
\author[6,21,22]{Pearu Peterson}
\author[23,24]{Pierre G\'erard-Marchant}
\author[25]{Kevin Sheppard}
\author[26]{Tyler Reddy}
\author[4]{Warren Weckesser}
\author[6]{Hameer Abbasi}
\author[27]{Christoph Gohlke}
\author[6]{Travis E. Oliphant}
\affil[1]{Independent Researcher, Logan, Utah, USA}
\affil[2]{Brain Imaging Center, University of California, Berkeley, Berkeley, CA, USA}
\affil[3]{Division of Biostatistics, University of California, Berkeley, Berkeley, CA, USA}
\affil[4]{Berkeley Institute for Data Science, University of California, Berkeley, Berkeley, CA, USA}
\affil[5]{Applied Mathematics, Stellenbosch University, Stellenbosch, South Africa}
\affil[6]{Quansight LLC, Austin, TX, USA}
\affil[7]{Department of Physics, University of Jyv\"askyl\"a, Jyv\"askyl\"a, Finland}
\affil[8]{Nanoscience Center, University of Jyv\"askyl\"a, Jyv\"askyl\"a, Finland}
\affil[9]{Mercari JP, Tokyo, Japan}
\affil[10]{Department of Engineering, University of Cambridge, Cambridge, UK}
\affil[11]{Independent Researcher, Karlsruhe, Germany}
\affil[12]{Independent Researcher, Berkeley, CA, USA}
\affil[13]{Enthought, Inc., Austin, TX, USA}
\affil[14]{Google Research, Mountain View, CA, USA}
\affil[15]{Department of Astronomy \& Astrophysics, University of Toronto, Toronto, ON, Canada}
\affil[16]{School of Psychology, University of Birmingham, Edgbaston, Birmigham, UK}
\affil[17]{Department of Physics, Temple University, Philadelphia, PA, USA}
\affil[18]{Google, Zurich, Switzerland}
\affil[19]{Department of Physics and Astronomy, The University of British Columbia, Vancouver, BC, Canada}
\affil[20]{Amazon, Seattle, Washington, USA}
\affil[21]{Independent Researcher, Saue, Estonia}
\affil[22]{Department of Mechanics and Applied Mathematics, Institute of Cybernetics at Tallinn Technical University, Tallinn, Estonia}
\affil[23]{Department of Biological and Agricultural Engineering, University of Georgia, Athens, GA}
\affil[24]{France-IX Services, Paris, France}
\affil[25]{Department of Economics, University of Oxford, Oxford, UK}
\affil[26]{CCS-7, Los Alamos National Laboratory, Los Alamos, NM, USA}
\affil[27]{Laboratory for Fluorescence Dynamics, Biomedical Engineering Department, University of California, Irvine, Irvine, CA, USA}
\affil[*]{millman@berkeley.edu, stefanv@berkeley.edu, ralf.gommers@gmail.com}